---
title: "Axon guidance molecule expression in developing SC"
author: "James Choi"
date: "Last compiled: `r Sys.Date()`"
format:
  html:
    code-fold: false
    toc: true
    toc-depth: 3
    number-sections: true
    link-external-newwindow: true
    fig-align: center
editor: source
execute: 
  cache: true
  warning: false
  message: false
project:
  execute-dir: project
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'D:/MiamiProject/Park_snSeq_SuperiorColliculus/')
```

## Background

Superior colliculus is innervated by RGC axons during the early post-natal period i.e. between E19 and P8. Innervation by these projections have previously been described to be sequential - different anatomical areas of SC are innervated at different time (e.g. anterior vs posterior). 

Hypothesis: Cells of SC express axon guidance molecules in time-dependent manner that coincides with RGC projection innervation.

```{r directories}
results_out <- 'results/axon-guidance-molecules/'
dir.create(path = results_out)
```


## Setup

```{r libraries}
library(Seurat)
library(ggplot2)
library(dplyr)
library(KEGGREST)
library(ComplexHeatmap)
library(patchwork)
```

```{r params}
my_feat_cols <- c('grey90','skyblue','royalblue','darkorchid','violetred','red')
str_wrap_factor <- function(x, ...) {
  levels(x) <- stringr::str_wrap(levels(x), ...)
  x
}
shuffle_rows <- function(df) {
  df[sample(1:nrow(df), size = nrow(df), replace = FALSE),]
}
str_wrap_factor <- function(x, ...) {
  levels(x) <- stringr::str_wrap(levels(x), ...)
  return(x)
}
time.cols <- c("#dd4539","#d9b123","#68c545","#514fd1")
celltype.cols <- RColorBrewer::brewer.pal(n = 12, name = 'Paired')
celltype.cols[11] <- 'gold'
neuron.cols <- c("#c2444c","#e14327","#be6231","#de8d26","#debc22","#b99f3e","#a6c336","#6f9a3e","#63d135","#59c251","#4db873","#36dbbc","#7b85dc","#6074eb","#554ea8","#6934c4","#9253ea","#542298","#ac69d3","#ce4ce2","#d980cc","#cc4bb5","#df35bd","#92337c","#d54681")
```

```{r load-data}
sc <- readRDS(file = 'data/sc.rds')
neuron <- readRDS(file = 'data/neuron.rds')
```

```{r umaps, fig.height=4.5, fig.width=10}
p1 <- DimPlot(sc, label = TRUE, group.by = 'celltype') +
  xlab(label = 'UMAP1') +
  ylab(label = 'UMAP2') +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = 'none')
p2 <- DimPlot(sc, group.by = 'time', split.by = 'time', ncol = 2) +
  xlab(label = 'UMAP1') +
  ylab(label = 'UMAP2') +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = 'none')
(p1 | p2) + patchwork::plot_layout()
```


## Guidance molecule gene set

First we query expression of the axon guidance molecules in all cell-types.


### GeneCopoeia ExProfile Array

From https://www.genecopoeia.com/product/pathway-gene-qpcr-arrays/

Gene list #013 (QG013): ExProfile™ Human Extracellular Matrix and Adhesion Molecules Related Gene qPCR Array

```{r qg.genes}
# This was written manually because reading the PDF was a pain.
qg.genes <- c('SPG7','FN1','CD44','HPRT1','VTN','VCAM1','TNC','CLEC3B','TIMP3','TIMP2','TIMP1','THBS3','THBS2','THBS1','TGFBI','SPP1','SPG7','SPARC','SGCE','SELP','SELL','SELE','PECAM1','MMP9','MMP8','MMP7','MMP3','MMP16','MMP15','MMP14','MMP13','MMP12','MMP11','MMP10','MMP1','LAMB1','LAMA3','LAMA2','LAMA1','KAL1','ITGB5','ITGB3','ITGB3','ITGB2','ITGB1','ITGAV','ITGAM','ITGAL','ITGA8','ITGA7','ITGA6','ITGA5','ITGA4','ITGA2','ITGA1','ICAM1','CTNND1','CTNNB1','CTNNA1','CTGF','VCAN','COL8A1','COL6A2','COL6A1','COL5A1','COL4A2','COL1A1','COL16A1','COL15A1','COL14A1','COL12A1','COL11A1','CNTN1','ADAMTS8','ADAMTS1','COL7A1','COL12A1','CDH1','ECM1','HAS1','ITGA3','GAPDH','ACTB','B2M','RPL13A','HPRT1','RN18S1')
# Mouse ortholog name conversion
qg.genes <- plyr::mapvalues(
  x = qg.genes, 
  from = c('HPRT1', 'MMP1'),
  to = c('HPRT', 'MMP1B')
)
qg.genes <- unique(tools::toTitleCase(text = tolower(qg.genes)))
qg.genes.missing <- qg.genes[!qg.genes %in% rownames(sc)]
qg.genes <- qg.genes[qg.genes %in% rownames(sc)]
```

### KEGG pathway: axon guidance

```{r kegg-genes}
kegg.guidance.entry <- keggGet(dbentries = 'mmu04360')
kegg.guidance.genes <- kegg.guidance.entry[[1]]$GENE[seq(0,length(kegg.guidance.entry[[1]]$GENE),2)]
kegg.guidance.genes <- gsub("\\;.*","", kegg.guidance.genes)
kegg.guidance.genes.missing <- kegg.guidance.genes[!kegg.guidance.genes %in% rownames(sc)]
kegg.guidance.genes <- kegg.guidance.genes[kegg.guidance.genes %in% rownames(sc)]
```

### Cadherins/Protocadherins

```{r}
cadherins <- c("CDH1", "CDH2", "CDH12", "CDH3", "DSG1", "DSG2", "DSG3", "DSG4", "DSC1", "DSC2", "DSC3", "PCDH1", "PCDH7", "PCDH8", "PCDH9", "PCDH10", "PCDH11X,", "PCDH11Y", "PCDH12", "PCDH15", "PCDH17", "PCDH18", "PCDH19", "PCDH20", "PCDHA1", "PCDHA2", "PCDHA3", "PCDHA4", "PCDHA5", "PCDHA6", "PCDHA7", "PCDHA8", "PCDHA9", "PCDHA10", "PCDHA11", "PCDHA12", "PCDHA13", "PCDHAC1", "PCDHAC2", "PCDHB1", "PCDHB2", "PCDHB3", "PCDHB4", "PCDHB5", "PCDHB6", "PCDHB7", "PCDHB8", "PCDHB9", "PCDHB10", "PCDHB11", "PCDHB12", "PCDHB13", "PCDHB14", "PCDHB15", "PCDHB16", "PCDHB17", "PCDHB18", "PCDHGA1", "PCDHGA2", "PCDHGA3", "PCDHGA4", "PCDHGA5", "PCDHGA6", "PCDHGA7", "PCDHGA8", "PCDHGA9", "PCDHGA10", "PCDHGA11", "PCDHGA12", "PCDHGB1", "PCDHGB2", "PCDHGB3", "PCDHGB4", "PCDHGB5", "PCDHGB6", "PCDHGB7", "PCDHGC3", "PCDHGC4", "PCDHGC5", "FAT", "FAT2", "FAT4", "CDH4", "CDH5", "CDH6", "CDH7", "CDH8", "CDH9", "CDH10", "CDH11", "CDH13", "CDH15", "CDH16", "CDH17", "CDH18", "CDH19", "CDH20", "CDH23", "CDH22", "CDH24", "CDH26", "CDH28", "CELSR1", "CELSR2", "CELSR3", "CLSTN1", "CLSTN2", "CLSTN3", "DCHS1", "DCHS2", "LOC389118", "PCLKC", "RESDA1", "RET")
cadherins <- tools::toTitleCase(tolower(cadherins))
cadherins <- cadherins[which(cadherins %in% rownames(sc[['RNA']]@counts))]
cadherin.avg.pct <- sparseMatrixStats::rowMeans2(sc[['RNA']]@counts[cadherins,] > 1)
cadherins.detected <- cadherins[which(cadherin.avg.pct > 0.001)]
```

## Plotting expression: QG013 gene set

### Dot plot expression across all cell-types

```{r qg0130-dot-plot, fig.height=13, fig.width=8}
Idents(sc) <- 'celltype'
DefaultAssay(sc) <- 'RNA'
tmp <- DotPlot(sc, assay = 'RNA', features = qg.genes, split.by = 'time', cols = rep('red',4))
plot.data <- tmp$data
plot.data$celltype <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 1)
plot.data$celltype = factor(x = plot.data$celltype, levels = levels(sc$celltype))
plot.data$time <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 2)
plot.data$time = factor(plot.data$time, levels = levels(sc$time))
plot.data = plot.data %>% 
  group_by(features.plot) %>% 
  mutate(avg.exp = scale(avg.exp)) %>%
  ungroup()
cutoffs = quantile(plot.data$avg.exp, c(0.025, 0.975), na.rm = TRUE)
plot.data = plot.data %>% 
  mutate(avg.exp = case_when(
    avg.exp < cutoffs[1] ~ cutoffs[1],
    avg.exp > cutoffs[2] ~ cutoffs[2],
    TRUE ~ avg.exp
  ))
p1 <- plot.data %>% 
  ggplot() + 
  geom_point(mapping = aes(x = time, 
                           y = features.plot,
                           size = pct.exp,
                           fill = avg.exp),
             pch = 21) + 
  facet_grid(. ~ str_wrap_factor(celltype, width = 10)) +
  scale_size(name = 'Percent\nexpression',
               range = c(0, 8)) +
  scale_fill_gradient2(
    low = 'blue2', 
    high = 'red2', 
    name = 'Average\nscaled\nexpression'
  ) +  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 16, angle = 65, hjust = 1),
        axis.text.y = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 14)) +
  guides(fill = guide_colorbar(frame.colour = 'black', 
                               ticks.colour = 'black'),
         size = guide_legend(override.aes = list(fill = 'black')))
ggsave(filename = paste0(results_out, 'QG013-guidance-molecules_dotplot.tiff'), plot = p1, height = 18, width = 15.5, device = 'tiff')
p1
```

The QG013 dot plot some cell-type specificity of certain molecules:


* Tnc is expressed mostly by astrocytes and its expression decreases over time.
* Among the MMPs, MMP15 and MMP16 are highest expressed by Oligo lineage cells with decreases over time.
* Among the integrins, Itga6, Itga7, and Itgb5 show astrocyte specificity. Itga6 increases in astrocytes over time.
* Laminins show specificity: 
    + Lama1 is restricted to fibroblasts/pericytes.
    + Lama2 expressed by astrocytes and fibroblasts/pericytes with increasing trend over time. Interesting that mature Oligos express during/after myelination.
    + Lamb1 expression in fibroblasts/pericytes decreases over time.
* Many of the collagens are specific to fibroblasts/pericytes and show trend to decrease over time.
* Fibronectin and vitronectin show similar pattern to collagens but with more convincing decreases over time.


```{r}
# dat <- FetchData(sc, vars = c(qg.genes, 'celltype', 'time'), slot = 'data') %>% 
#   mutate(across(.cols = where(is.numeric), .fns = scale))
# dat_celltype <- dat$celltype
# dat_time <- dat$time
# is_numeric <- function(x) return(all(!is.na(x) & is.numeric(x)))
# dat <- dat %>% select(where(is_numeric)) %>% as.matrix()
# Heatmap(
#   matrix = dat,
#   clustering_method_rows = 'ward.D2',
#   clustering_method_columns = 'ward.D2',
#   show_row_names = FALSE,
#   show_column_names = TRUE,
# )
```

Heatmap approach to identifying QG-gene expression patterns gives inconclusive result/visualization (above).

### UMAP expression across time-points

```{r tnc-umap, collapse=TRUE, fig.height=2.5, fig.width=10, fig.cap='Tnc expression'}
genes <- c('Tnc')
p.tnc <- FetchData(sc, c('time', genes, 'UMAP_1','UMAP_2')) %>% 
  reshape2::melt(id.vars = c('UMAP_1','UMAP_2','time')) %>% 
  arrange(value) %>% 
  ggplot() +
  geom_point(aes(x = UMAP_1, y = UMAP_2, color = value), size = 0.1) +
  facet_grid(variable ~ time) +
  scale_color_gradientn(colors = my_feat_cols) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 12)) +
  guides(color = guide_colorbar(frame.colour = 'black', ticks.colour = 'black', title = 'Scaled\nexpression'))
ggsave(filename = paste0(results_out, 'umap-tnc.tiff'), plot = p.tnc, height = 2.65, width = 10, device = 'tiff', dpi = 320)
p.tnc
```

```{r mmp-umap, collapse=TRUE, fig.height=5, fig.width=10, fig.cap='MMP expression'}
genes <- c('Mmp15','Mmp16')
p.mmp <- FetchData(sc, c('time', genes, 'UMAP_1','UMAP_2')) %>% 
  reshape2::melt(id.vars = c('UMAP_1','UMAP_2','time')) %>% 
  arrange(value) %>% 
  ggplot() +
  geom_point(aes(x = UMAP_1, y = UMAP_2, color = value), size = 0.1) +
  facet_grid(variable ~ time) +
  scale_color_gradientn(colors = my_feat_cols) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 12)) +
  guides(color = guide_colorbar(frame.colour = 'black', ticks.colour = 'black', title = 'Scaled\nexpression'))
ggsave(filename = paste0(results_out, 'umap-mmp.tiff'), plot = p.mmp, height = 4.85, width = 10, device = 'tiff', dpi = 320)
p.mmp
```

```{r integrin-umap, collapse=TRUE, fig.height=7.5, fig.width=10, fig.cap='Integrin expression'}
genes <- c('Itga6', 'Itga7', 'Itgb5')
p.integrin <- FetchData(sc, c('time', genes, 'UMAP_1','UMAP_2')) %>% 
  reshape2::melt(id.vars = c('UMAP_1','UMAP_2','time')) %>% 
  arrange(value) %>% 
  ggplot() +
  geom_point(aes(x = UMAP_1, y = UMAP_2, color = value), size = 0.1) +
  facet_grid(variable ~ time) +
  scale_color_gradientn(colors = my_feat_cols) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 12)) +
  guides(color = guide_colorbar(frame.colour = 'black', ticks.colour = 'black', title = 'Scaled\nexpression'))
ggsave(filename = paste0(results_out, 'umap-integrin.tiff'), plot = p.integrin, height = 7, width = 10, device = 'tiff', dpi = 320)
p.integrin
```

```{r laminin-umap, collapse=TRUE, fig.height=7.5, fig.width=10, fig.cap='Laminin expression'}
genes <- c('Lama1','Lama2','Lamb1')
p.laminin <- FetchData(sc, c('time', genes, 'UMAP_1','UMAP_2')) %>% 
  reshape2::melt(id.vars = c('UMAP_1','UMAP_2','time')) %>% 
  arrange(value) %>% 
  ggplot() +
  geom_point(aes(x = UMAP_1, y = UMAP_2, color = value), size = 0.1) +
  facet_grid(variable ~ time) +
  scale_color_gradientn(colors = my_feat_cols) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 12)) +
  guides(color = guide_colorbar(frame.colour = 'black', ticks.colour = 'black', title = 'Scaled\nexpression'))
ggsave(filename = paste0(results_out, 'umap-laminin.tiff'), plot = p.laminin, height = 7, width = 10, device = 'tiff', dpi = 320)
p.laminin
```

```{r ecm-umap, collapse=TRUE, fig.height=7.5, fig.width=10, fig.cap='Other EMC molecule expression'}
genes <- c('Col1a1','Vtn','Fn1')
p.ecm <- FetchData(sc, c('time', genes, 'UMAP_1','UMAP_2')) %>% 
  reshape2::melt(id.vars = c('UMAP_1','UMAP_2','time')) %>% 
  arrange(value) %>% 
  ggplot() +
  geom_point(aes(x = UMAP_1, y = UMAP_2, color = value), size = 0.1) +
  facet_grid(variable ~ time) +
  scale_color_gradientn(colors = my_feat_cols) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        strip.text = element_text(size = 12)) +
  guides(color = guide_colorbar(frame.colour = 'black', ticks.colour = 'black', title = 'Scaled\nexpression'))
ggsave(filename = paste0(results_out, 'umap-ecm.tiff'), plot = p.ecm, height = 7, width = 10, device = 'tiff', dpi = 320)
p.ecm
```


## Plotting expression: KEGG pathway: axon guidance

### Dot plot expression across all cell-types

```{r kegg-dot-plot, fig.height=24, fig.width=8}
DefaultAssay(sc) <- 'RNA'
tmp <- DotPlot(sc, assay = 'RNA', features = sort(kegg.guidance.genes, decreasing = TRUE), split.by = 'time', cols = rep('red',4))
plot.data = tmp$data
plot.data$celltype <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 1)
plot.data$celltype = factor(x = plot.data$celltype, levels = levels(sc$celltype))
plot.data$time <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 2)
plot.data$time = factor(plot.data$time, levels = levels(sc$time))
plot.data = plot.data %>% 
  group_by(features.plot) %>% 
  mutate(avg.exp = scale(avg.exp)) %>%
  ungroup()
cutoffs = quantile(plot.data$avg.exp, c(0.025, 0.975), na.rm = TRUE)
plot.data = plot.data %>% 
  mutate(avg.exp = case_when(
    avg.exp < cutoffs[1] ~ cutoffs[1],
    avg.exp > cutoffs[2] ~ cutoffs[2],
    TRUE ~ avg.exp
  ))
p2 <- plot.data %>% 
  ggplot() + 
  geom_point(mapping = aes(x = time, 
                           y = features.plot,
                           size = pct.exp,
                           fill = avg.exp),
             pch = 21) + 
  facet_grid(. ~ str_wrap_factor(celltype, width = 10)) +
  scale_size(name = 'Percent\nexpression',
               range = c(0, 8)) +
  scale_fill_gradient2(
    low = 'blue2', 
    high = 'red2', 
    name = 'Average\nscaled\nexpression'
  ) +  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 16, angle = 65, hjust = 1),
        axis.text.y = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 14)) +
  guides(fill = guide_colorbar(frame.colour = 'black', 
                               ticks.colour = 'black'),
         size = guide_legend(override.aes = list(fill = 'black')))
ggsave(filename = paste0(results_out, 'KEGG-guidance-molecules_dotplot.tiff'), plot = p2, height = 34, width = 16, device = 'tiff')
p2
```


### Topographic mapping associated genes

```{r kegg-genes}
kegg.guidance.entry <- keggGet(dbentries = 'mmu04360')
kegg.guidance.genes <- kegg.guidance.entry[[1]]$GENE[seq(0,length(kegg.guidance.entry[[1]]$GENE),2)]
kegg.guidance.genes <- gsub("\\;.*","", kegg.guidance.genes)
kegg.guidance.genes.missing <- kegg.guidance.genes[!kegg.guidance.genes %in% rownames(sc)]
kegg.guidance.genes <- kegg.guidance.genes[kegg.guidance.genes %in% rownames(sc)]
ephrin.genes = sort(grep(pattern = 'Eph|Efn', x = kegg.guidance.genes, value = TRUE), decreasing = TRUE)
sema.plxn.genes = sort(grep(pattern = 'Sema|Plxn', x = kegg.guidance.genes, value = TRUE), decreasing = TRUE)
slit.robo.genes = sort(grep(pattern = 'Robo|Slit', x = kegg.guidance.genes, value = TRUE), decreasing = TRUE)
topo.genes = data.frame(
  gene = c(ephrin.genes, sema.plxn.genes, slit.robo.genes),
  family = c(rep('Eph Receptor - Ephrin', length(ephrin.genes)),
             rep('Semaphorin - Plexin', length(sema.plxn.genes)),
             rep('Slit - Robo', length(slit.robo.genes)))
)
```

#### Dot plot by molecule class (all celltypes)

```{r}
Idents(sc) = 'celltype'
tmp <- DotPlot(sc, assay = 'RNA', features = topo.genes$gene, split.by = 'time', cols = rep('red',4))
plot.data <- tmp$data
plot.data$celltype <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 1)
plot.data$celltype = factor(x = plot.data$celltype, levels = levels(sc$celltype))
plot.data$celltype = plyr::mapvalues(
  x = plot.data$celltype,
  from = 'Vascular Leptomeningeal Cell',
  to = 'Vascular Lepto-meningeal Cell'
)
plot.data$time <- sapply(strsplit(as.character(plot.data$id), '_'), `[`, 2)
plot.data$time = factor(plot.data$time, levels = levels(sc$time))
plot.data = plot.data %>% 
  group_by(features.plot) %>% 
  mutate(avg.exp = scale(avg.exp)) %>%
  ungroup() %>% 
  mutate(Family = plyr::mapvalues(
    x = features.plot,
    from = topo.genes$gene,
    to = topo.genes$family
  ))
cutoffs = quantile(plot.data$avg.exp, c(0.025, 0.975), na.rm = TRUE)
plot.data = plot.data %>% 
  mutate(avg.exp = case_when(
    avg.exp < cutoffs[1] ~ cutoffs[1],
    avg.exp > cutoffs[2] ~ cutoffs[2],
    TRUE ~ avg.exp
  ))
p3 <- plot.data %>% 
  ggplot() + 
  geom_point(mapping = aes(x = time, 
                           y = features.plot,
                           size = pct.exp,
                           fill = avg.exp),
             pch = 21) + 
  facet_grid(Family ~ str_wrap_factor(celltype, width = 10),
             scales = 'free_y',
             space = 'free_y') +
  scale_size(name = 'Percent\nexpression',
               range = c(0, 8)) +
  scale_fill_gradient2(
    low = 'blue2', 
    high = 'red2', 
    name = 'Average\nscaled\nexpression'
  ) +  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 16, angle = 65, hjust = 1),
        axis.text.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 18),
        strip.text.y = element_text(size = 20),
        legend.position = 'bottom') +
  guides(fill = guide_colorbar(frame.colour = 'black', 
                               ticks.colour = 'black'),
         size = guide_legend(override.aes = list(fill = 'black')))
ggsave(filename = paste0(results_out, 'topographic-mapping-molecules_dotplot.tiff'), plot = p3, height = 16.5, width = 13, device = 'tiff')
p3
```


#### Dot plot by molecule class (neuron subtypes)

```{r}
Idents(neuron) = 'subtype'
# detected.topo.genes = sparseMatrixStats::rowSums2(
#   x = neuron@assays$RNA@counts[topo.genes$gene,] > 0
# ) > 50 # because smallest neuron subtype has 111 cells, so we set about half
# detected.topo.genes = topo.genes$gene[detected.topo.genes]
tmp <- DotPlot(neuron, assay = 'RNA', features = topo.genes$gene)
plot.data <- tmp$data
plot.data = plot.data %>% 
  mutate(Family = plyr::mapvalues(
    x = features.plot,
    from = topo.genes$gene,
    to = topo.genes$family
  ))
cutoffs = quantile(plot.data$avg.exp.scaled, c(0.025, 0.975), na.rm = TRUE)
plot.data = plot.data %>% 
  mutate(avg.exp.scaled = case_when(
    avg.exp.scaled < cutoffs[1] ~ cutoffs[1],
    avg.exp.scaled > cutoffs[2] ~ cutoffs[2],
    TRUE ~ avg.exp.scaled
  ))
p3 <- plot.data %>% 
  ggplot() + 
  geom_point(
    mapping = aes(x = id, y = features.plot, size = pct.exp, fill = avg.exp.scaled),
    pch = 21
  ) + 
  facet_grid(Family ~ ., space = 'free_y', drop = TRUE, scales = 'free_y') +
  scale_size(name = 'Percent\nexpression') +
  scale_fill_gradient2(
    low = 'blue2', 
    high = 'red2', 
    name = 'Average\nscaled\nexpression'
  ) + 
  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 12, angle = 65, hjust = 1),
        axis.text.y = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14),
        legend.position = 'bottom') +
  guides(fill = guide_colorbar(frame.colour = 'black', 
                               ticks.colour = 'black'),
         size = guide_legend(override.aes = list(fill = 'black')))
p3
# ggsave(filename = paste0(results_out, 'topographic-mapping-molecules_dotplot.tiff'), plot = p3, height = 16.5, width = 13, device = 'tiff')

```

### Guidance molecule expression at subtype-level

```{r}
sc$subtype <- colnames(sc)
sc$subtype <- plyr::mapvalues(
  x = sc$subtype,
  from = colnames(neuron),
  to = paste(neuron$celltype_adjusted, neuron$subtype, sep = '_'),
  warn_missing = FALSE
)
sc$subtype[!colnames(sc) %in% colnames(neuron)] <- plyr::mapvalues(
  x = sc$subtype[!colnames(sc) %in% colnames(neuron)],
  from = colnames(sc)[!colnames(sc) %in% colnames(neuron)],
  to = as.character(sc$celltype[!colnames(sc) %in% colnames(neuron)]),
  warn_missing = FALSE
)
```

#### Heatmap across celltypes

```{r all-cells-subtype-heatmap}
Idents(sc) <- 'subtype'
tmp <- DotPlot(sc, features = kegg.guidance.genes, assay = 'RNA')
plot.data <- tmp$data
tmp <- plot.data %>% 
  filter(!is.na(avg.exp.scaled)) %>% 
  reshape2::acast(formula = id ~ features.plot, value.var = 'avg.exp.scaled')
subtype.hclust <- hclust(d = dist(x = tmp), method = 'ward.D2')
gene.hclust <- hclust(d = dist(x = t(tmp)), method = 'ward.D2')

tiff(filename = paste0(results_out, 'kegg-guidance-molecules_subtype-heatmap.tiff'), height = 6, width = 12, res = 440, units = 'in')
draw(ComplexHeatmap::Heatmap(
  matrix = tmp,
  cluster_rows = dendsort::dendsort(subtype.hclust),
  cluster_columns = dendsort::dendsort(gene.hclust),
  # show_row_names = FALSE,
  # show_column_names = TRUE,
  show_column_names = FALSE
))
dev.off()

tiff(filename = paste0(results_out, 'kegg-guidance-molecules_subtype-heatmap_labeled.tiff'), height = 7.5, width = 26, res = 320, units = 'in')
draw(ComplexHeatmap::Heatmap(
  matrix = tmp,
  cluster_rows = dendsort::dendsort(subtype.hclust),
  cluster_columns = dendsort::dendsort(gene.hclust),
  # show_row_names = FALSE,
  show_column_names = TRUE,
  # show_column_names = FALSE
))
dev.off()
```

```{r}
# DefaultAssay(sc) <- 'RNA'
# dat <- FetchData(
#   object = sc, 
#   vars = c(kegg.guidance.genes, 'celltype', 'time'),
#   slot = 'data'
# ) %>%
#   mutate(across(.cols = where(is.numeric), .fns = scale))
# dat_celltype <- dat$celltype
# dat_time <- dat$time
# is_numeric <- function(x) return(all(!is.na(x) & is.numeric(x)))
# dat <- dat %>% select(where(is_numeric)) %>% as.matrix()
# Heatmap(
#   matrix = dat,
#   clustering_method_rows = 'ward.D2',
#   clustering_method_columns = 'ward.D2',
#   show_row_names = FALSE,
#   show_column_names = TRUE,
# )
```

Heatmap with KEGG guidance genes shows clustering of guidance molecule expression non-neuronal cells.

#### Heatmap across neuron subtypes

```{r neurons-subtype-heatmap}
neuron$subtype.2 <- paste(neuron$celltype_adjusted, neuron$subtype, sep = '_')
Idents(neuron) <- 'subtype.2'
plot.data <- DotPlot(neuron, features = kegg.guidance.genes, assay = 'RNA')
plot.data <- plot.data$data
tmp <- plot.data %>% 
  filter(!is.na(avg.exp.scaled)) %>% 
  reshape2::acast(formula = id ~ features.plot, value.var = 'avg.exp.scaled')
subtype.hclust <- hclust(d = dist(x = tmp), method = 'ward.D2')
gene.hclust <- hclust(d = dist(x = t(tmp)), method = 'ward.D2')

tiff(filename = paste0(results_out, 'kegg-guidance-molecules_neurons-subtype-heatmap.tiff'), height = 5.5, width = 9, res = 320, units = 'in')
draw(ComplexHeatmap::Heatmap(
  matrix = tmp,
  cluster_rows = dendsort::dendsort(subtype.hclust),
  cluster_columns = dendsort::dendsort(gene.hclust),
  # show_row_names = FALSE,
  show_column_names = FALSE,
  heatmap_legend_param = list(
    title = 'Scaled expression'
  )
))
dev.off()

{
  tiff(filename = paste0(results_out, 'kegg-guidance-molecules_neurons-subtype-heatmap_labeled.tiff'), height = 6.5, width = 28, res = 320, units = 'in')
  draw(ComplexHeatmap::Heatmap(
    matrix = tmp,
    cluster_rows = dendsort::dendsort(subtype.hclust),
    cluster_columns = dendsort::dendsort(gene.hclust),
    # column_names_rot = 65,
    # show_row_names = FALSE,
    show_column_names = TRUE,
    heatmap_legend_param = list(
      title = 'Scaled expression'
    )
  ))
  dev.off()
}
```

Heatmap of KEGG guidance genes in neuronal subtypes show no distinct clusters of patterns of expression.


## Clustering all cells on guidance molecule expression

```{r}
DefaultAssay(sc) <- 'RNA'
guidance.genes <- unique(c(qg.genes, kegg.guidance.genes, cadherins.detected))
axon_guidance <- SplitObject(sc, split.by = 'orig.ident')
axon_guidance <- lapply(
  X = axon_guidance,
  FUN = NormalizeData
)
axon_guidance <- FindIntegrationAnchors(
  object.list = axon_guidance,
  anchor.features = guidance.genes
)
axon_guidance <- IntegrateData(
  anchorset = axon_guidance,
  features = guidance.genes
)
DefaultAssay(axon_guidance) <- 'integrated'
axon_guidance <- axon_guidance %>% 
  ScaleData() %>% 
  RunPCA()
ElbowPlot(axon_guidance, ndims = 40)
axon_guidance <- axon_guidance %>% 
  FindNeighbors(dims = 1:10) %>% 
  RunUMAP(dims = 1:10) %>% 
  FindClusters(resolution = 0.1)
axon_guidance$celltype = factor(axon_guidance$celltype, levels = levels(sc$celltype))
axon_guidance$time = factor(axon_guidance$time, levels = levels(sc$time))
```

### Multi-Dimensional Scaling (MDS)

```{r}
DefaultAssay(axon_guidance) <- 'RNA'
meta_vars <- c('celltype','time')
dat <- FetchData(
  object = axon_guidance,
  vars =  c(meta_vars, guidance.genes)
)
dat <- dat %>% 
  group_by(across(all_of(meta_vars))) %>% 
  summarise(across(.cols = where(is.numeric), .fns = mean))
dat_meta <- dat[meta_vars]
dat_avg <- as.data.frame(dat[!colnames(dat) %in% meta_vars])
rownames(dat_avg) <- paste(dat_meta$celltype, dat_meta$time, sep = '.')
dat_dist <- dist(x = dat_avg, method = 'euclidean')
mds <- as.data.frame(cmdscale(d = dat_dist, k = 7))
colnames(mds) <- paste('MDS', 1:7, sep = '_')
mds_dat <- cbind(mds, dat_meta)
mds.celltype = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = celltype), 
             color = 'black', pch = 21, size = 3) + 
  scale_fill_manual(values = celltype.cols) + 
  labs(title = 'MDS colored by celltype') +
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        legend.title = element_blank(),
        plot.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA)) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
mds.time = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = time), 
             color = 'black', pch = 21, size = 3) + 
  labs(title = 'MDS colored by time-points') +
  scale_fill_manual(values = time.cols) + 
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        plot.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA)) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
mds = mds.celltype | mds.time
ggsave(filename = paste0(results_out, 'axon-guidance-mds-plot.tiff'), plot = mds, device = 'tiff', height = 3.75, width = 10, dpi = 320)
```

### PCA

```{r}
pca.axes <- ggplot(data.frame(x=10,y=10),aes(x=x,y,y)) +
  geom_point() +
  xlim(c(0,1)) + ylim(c(0,1)) +
  theme_classic() +
  xlab('PC 1') + ylab('PC 2') +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 10),
        axis.line = element_line(
          color = 'black',
          arrow = arrow(angle = 15, length = unit(.35, 'cm'), type = 'closed')
        ),
        panel.background = element_rect(fill = 'transparent', color = NA),
        panel.border = element_rect(fill = 'transparent', color = NA),
  )
layout <- c(
  patchwork::area(t = 11, l = 1, b = 11, r = 1),
  patchwork::area(t = 1, l = 1, b = 11, r = 11)
)
pca.celltype <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'celltype', shuffle = TRUE, label = TRUE, repel = TRUE, label.size = 4) +
  scale_color_manual(values = celltype.cols) +
  labs(title = 'PCA colored by celltype') +
  theme_void() +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 18))
pca.celltype = pca.axes + pca.celltype + plot_layout(design = layout)
pca.time <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'time', shuffle = TRUE) +
  labs(title = 'PCA colored by time-point') +
  scale_color_manual(values = time.cols) +
  theme_void() +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 18))
pca.time = pca.axes + pca.time + plot_layout(design = layout)
pca.axon = pca.celltype | pca.time
ggsave(filename = paste0(results_out, 'axon-guidance-pca-plot.tiff'), plot = pca.axon, device = 'tiff', height = 4.5, width = 11, dpi = 320)
```

### Summary

```{r}
pca.celltype <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'celltype', shuffle = TRUE, label = TRUE, repel = TRUE, label.size = 4, pt.size = 1) +
  scale_color_manual(values = celltype.cols) +
  labs(title = 'PCA colored by celltype') +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 16, hjust = 0, face = 'plain'),
        legend.position = 'none',
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  coord_fixed()
mds.celltype = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = celltype), 
             color = 'black', pch = 21, size = 3.5) + 
  scale_fill_manual(values = celltype.cols) + 
  labs(title = 'MDS colored by celltype') +
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        legend.title = element_blank(),
        plot.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
mds.timepoint = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = time), 
             color = 'black', pch = 21, size = 3.5) + 
  scale_fill_manual(values = time.cols) + 
  labs(title = 'MDS colored by time-point') +
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        legend.title = element_blank(),
        plot.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
dimreduc.plot = pca.celltype | mds.celltype
ggsave(filename = paste0(results_out, 'axon-guidance-all-cells-dim-reduction-summary.tiff'), plot = dimreduc.plot, device = 'tiff', height = 4.5, width = 11, dpi = 320)
```

Dim reduction and clustering of all neurons using only guidance molecule genes shows some smaller clusters of distinct subsets, and some overlapping/distinct regions of neuronal ex/inh types but with no relation to timepoint.

## Clustering neurons on guidance molecule expression

```{r}
DefaultAssay(neuron) <- 'RNA'
guidance.genes <- unique(c(qg.genes, kegg.guidance.genes, cadherins.detected))
axon_guidance <- SplitObject(neuron, split.by = 'orig.ident')
axon_guidance <- lapply(
  X = axon_guidance,
  FUN = NormalizeData
)
axon_guidance <- FindIntegrationAnchors(
  object.list = axon_guidance,
  anchor.features = guidance.genes
)
axon_guidance <- IntegrateData(
  anchorset = axon_guidance,
  features = guidance.genes
)
DefaultAssay(axon_guidance) <- 'integrated'
axon_guidance <- axon_guidance %>% 
  ScaleData() %>% 
  RunPCA()
ElbowPlot(axon_guidance, ndims = 40)
axon_guidance <- axon_guidance %>% 
  FindNeighbors(dims = 1:10) %>% 
  RunUMAP(dims = 1:10) %>% 
  FindClusters(resolution = 0.1)
axon_guidance$subtype = factor(axon_guidance$subtype, levels = levels(neuron$subtype))
axon_guidance$time = factor(axon_guidance$time, levels = levels(neuron$time))
```

### Multi-Dimensional Scaling (MDS)

```{r}
DefaultAssay(axon_guidance) <- 'RNA'
meta_vars <- c('subtype')
dat <- FetchData(
  object = axon_guidance,
  vars =  c(meta_vars, guidance.genes)
)
dat <- dat %>% 
  group_by(across(all_of(meta_vars))) %>% 
  summarise(across(.cols = where(is.numeric), .fns = mean))
dat_meta <- dat[meta_vars]
dat_avg <- as.data.frame(dat[!colnames(dat) %in% meta_vars])
rownames(dat_avg) <- dat_meta$subtype
dat_dist <- dist(x = dat_avg, method = 'euclidean')
mds <- as.data.frame(cmdscale(d = dat_dist, k = 7))
colnames(mds) <- paste('MDS', 1:7, sep = '_')
mds_dat <- cbind(mds, dat_meta)
mds.subtype = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = subtype), 
             color = 'black', pch = 21, size = 3) + 
  scale_fill_manual(values = neuron.cols) + 
  labs(title = 'MDS colored by subtype') +
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        legend.title = element_blank(),
        plot.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA)) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
mds.subtype
ggsave(filename = paste0(results_out, 'axon-guidance-neuron-subtype-mds-plot.tiff'), plot = mds.subtype, device = 'tiff', height = 3.75, width = 5, dpi = 320)
```

### PCA

```{r}
pca.axes <- ggplot(data.frame(x=10,y=10),aes(x=x,y,y)) +
  geom_point() +
  xlim(c(0,1)) + ylim(c(0,1)) +
  theme_classic() +
  xlab('PC 1') + ylab('PC 2') +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 10),
        axis.line = element_line(
          color = 'black',
          arrow = arrow(angle = 15, length = unit(.35, 'cm'), type = 'closed')
        ),
        panel.background = element_rect(fill = 'transparent', color = NA),
        panel.border = element_rect(fill = 'transparent', color = NA),
  )
layout <- c(
  patchwork::area(t = 11, l = 1, b = 11, r = 1),
  patchwork::area(t = 1, l = 1, b = 11, r = 11)
)
pca.subtype <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'subtype', shuffle = TRUE, label = TRUE, repel = TRUE, label.size = 4) +
  scale_color_manual(values = neuron.cols) +
  labs(title = 'PCA colored by subtype') +
  theme_void() +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 18))
pca.subtype = pca.axes + pca.subtype + plot_layout(design = layout)
pca.time <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'time', shuffle = TRUE) +
  labs(title = 'PCA colored by time-point') +
  scale_color_manual(values = time.cols) +
  theme_void() +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 18))
pca.time = pca.axes + pca.time + plot_layout(design = layout)
pca.axon = pca.subtype | pca.time
ggsave(filename = paste0(results_out, 'axon-guidance-neuron-subtype-pca-plot.tiff'), plot = pca.axon, device = 'tiff', height = 4.5, width = 11, dpi = 320)
```

### Summary

```{r}
pca.subtype <- DimPlot(axon_guidance, reduction = 'pca', group.by = 'subtype', shuffle = TRUE, pt.size = 1) +
  scale_color_manual(values = neuron.cols) +
  labs(title = 'PCA colored by subtype') +
  theme(legend.text = element_text(size = 12),
        plot.title = element_text(size = 16, hjust = 0, face = 'plain'),
        legend.position = 'none',
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  coord_fixed()
mds.subtype = mds_dat %>% 
  shuffle_rows() %>% 
  ggplot(mapping = aes(x = MDS_1, y = MDS_2)) +
  geom_point(mapping = aes(fill = subtype), 
             color = 'black', pch = 21, size = 3.5) + 
  scale_fill_manual(values = neuron.cols) + 
  labs(title = 'MDS colored by subtype') +
  theme_bw() +
  theme(axis.title = element_text(size = 12),
        legend.title = element_blank(),
        plot.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        panel.background = element_rect(color = 'black', fill = NA),
        panel.border = element_rect(color = 'black', fill = NA),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  guides(fill = guide_legend(override.aes = list(size = 5))) +
  coord_fixed()
dimreduc.plot = pca.subtype | mds.subtype
ggsave(filename = paste0(results_out, 'axon-guidance-neuron-subtype-dim-reduction-summary.tiff'), plot = dimreduc.plot, device = 'tiff', height = 4.5, width = 11, dpi = 320)
```





```{r time-course, eval=FALSE}
# DefaultAssay(sc) <- 'RNA'
# tmp <- ScaleData(object = sc, features = union(kegg.guidance.genes, qg.genes))
# tmp <- FetchData(
#   object = sc, 
#   vars = c(union(kegg.guidance.genes, qg.genes), 'time', 'celltype'), 
#   slot = 'data'
# ) %>% 
#   mutate(across(where(is.numeric), scale)) %>%
#   group_by(celltype, time) %>% 
#   summarise(across(.cols = everything(), .fns = mean)) %>%
#   ungroup() %>% 
#   reshape2::melt(id.vars = c('celltype','time')) %>% 
#   mutate(obs_id = paste(celltype, variable, sep = '_')) %>% 
#   select(c('obs_id', 'time', 'value')) %>% 
#   reshape2::acast(formula = obs_id ~ time)
# tmp.dist <- dist(x = tmp)
# tmp.hclust <- hclust(d = tmp.dist, method = 'ward.D2')
# tmp.cluster <- cutree(tree = tmp.hclust, h = 10)
# cbind(as.data.frame(tmp), 'cluster' = tmp.cluster) %>% 
#   tibble::rownames_to_column('obs') %>% 
#   reshape2::melt(id.vars = c('obs', 'cluster')) %>% 
#   ggplot(mapping = aes(x = variable, y = value, group = obs, color = cluster)) +
#   geom_line() +
#   theme(legend.position = 'none')
```

## END

```{r}
sessionInfo()
```

